Cyclist video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda

### 2024-04-22 paper.qmd

Dangerous and criminal driving are significant problems that take many forms [@simon1996] [@corbett2010; @corbett2003] (Corbett 2003, 2010).
Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities (ref).
Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).

@@

However, it is a noted limitation that these are not precise geolocations but allow some broad analysis of the clustering of offences. 
XX THERE WAS MISSING LOCATION DATA FOR SOME MONTHS – SO IT WASN’T ‘LOCATION NOT AVAILABLE’ MISSINGS BUT LOOKED MORE LIKE IT WAS OMITTED ACCIDENTALLY OR CUT?

@@

```{r}
#| echo: false
#| include: false
d_all = read_csv(file_name)
d_all = d_all |>
  mutate(
    mode = tolower(mode),
    offence = tolower(offence),
    disposal = tolower(disposal)
  )
d_all |>
  head(3) |>
  knitr::kable()
```

@@

There are `r nrow(d_all)` records in the data, with increasing numbers of records over time (average n. records per month shown in @fig-time):

```{r}
#| label: fig-time
#| fig-cap: "Number of records per month in the opSnap dataset for West Yorkshire"
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "all")
d_offence = d_all |>
  opsnap:::filter_offence_nas()
d_offence_monthly = d_offence |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete_offence")
d = d_offence |>
  opsnap:::filter_location_nas()
d_complete_monthly = d |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete")
d_monthly = bind_rows(d_all_monthly, d_offence_monthly, d_complete_monthly)
d_monthly |>
  ggplot() +
  geom_line(aes(month, n, colour = records), alpha = 0.5, size = 2) +
  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire\nPolice Operation Snap data",
       x = "Date",
       y = "Number of records per month") +
  theme_minimal()
```

@@

The breakdown of records by mode of transport (of the observer) is shown below:

```{r}
d_all |>
  count(mode, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

The offence text strings are quite long, with the most common offences shown below:

```{r}
d_all |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent table excluding records with missing offence data is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

The equivalent for cyclists, with the least common offences categorised as 'other', is shown below:

```{r}
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  filter(mode == "cyclist") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  mutate(
    offence = ifelse(n < 20, "other", offence)
  ) |>
  group_by(offence) |>
  summarise(n = sum(n), n_hybrid = n()) |>
  arrange(n_hybrid, desc(n)) |>
  select(-n_hybrid) |>
  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
  knitr::kable()
```

In terms 'disposal', the most common values are shown below:

```{r}
d_all |>
  count(disposal, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  knitr::kable()
```

There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown below:

```{r}
#| echo: false
d |>
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d)) |>
  mutate(percent_records = scales::percent(percent_records)) |>
  arrange(desc(n)) |>
  head(10) |>
  knitr::kable()
```

>>>>

Dangerous and criminal driving are significant problems that take many forms [@simon1996; @corbett2003; @corbett2010].
Across Great Britain there were an estimated 1,711 fatalities resulting from a road traffic crash in 2022 and 135,480 causalities reported across all severities (ref).
Most injuries involve motor vehicle occupants, however when accounting for the distance travelled by road user groups, cyclists are over-represented in injury statistics (ref).

@@

@@

```{r}
#| echo: false
#| include: false
d_all = read_csv(file_name)
d_all = d_all |>
  mutate(
    mode = tolower(mode),
    offence = tolower(offence),
    disposal = tolower(disposal)
  )
d_all |>
  head(3) |>
  # Convert column names to title case:
  rename_all(snakecase::to_title_case) |>
  # snakecase::to_title_case
  knitr::kable()
```

@@

There are `r nrow(d_all)` records in the data, with increasing numbers of records over time, as illustrated in @fig-time:

```{r}
#| label: fig-time
#| fig-cap: "Number of records per month in the opSnap dataset for West Yorkshire"
d_all_monthly = d_all |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "all")
d_offence = d_all |>
  opsnap:::filter_offence_nas()
d_offence_monthly = d_offence |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete_offence")
d = d_offence |>
  opsnap:::filter_location_nas()
d_complete_monthly = d |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  group_by(month) |>
  summarise(n = n()) |>
  mutate(records = "complete")
d_monthly = bind_rows(d_all_monthly, d_complete_monthly)
d_monthly |>
  mutate(records = snakecase::to_title_case(records)) |> 
  rename_all(snakecase::to_title_case) |> 
  ggplot() +
  geom_line(aes(Month, N, colour = Records), alpha = 0.5, size = 2) +
  # geom_smooth(aes(month, n, colour = records), method = "lm", se = FALSE) +
  labs(title = "Number of records in West Yorkshire Police\nOperation Snap data",
       x = "Date",
       y = "Number of records per month") +
  theme_minimal()
```

@@

The breakdown of records by mode of transport (of the observer) is shown @tbl-mode.

```{r}
#| label: tbl-mode
#| tbl-cap: Number of OpSnap records by observer mode.
d_all |>
  count(mode, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  arrange(desc(n)) |>
  rename_all(snakecase::to_title_case) |>
  knitr::kable()
```

The offence text strings are quite long, with the most common offences shown in @tbl-offences:

```{r}
#| label: tbl-offences
#| tbl-cap: Number and percentages of OpSnap records by offence type.
d_all |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  arrange(desc(n)) |>
  head(10) |>
  rename_all(snakecase::to_title_case) |>
  knitr::kable()
```

<!-- The equivalent table excluding records with missing offence data is shown below: -->

```{r}
#| include: false
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  arrange(desc(n)) |>
  head(10) |>
  rename_all(snakecase::to_title_case) |>
  knitr::kable()
```

The equivalent for cyclists, with the least common offences categorised as 'other', is shown in @tbl-offences-cyclist-observer.

```{r}
#| label: tbl-offences-cyclist-observer
#| tbl-cap: "Number and percentages of OpSnap records, submitted by cyclists, by offence type."
d_all |>
  opsnap:::filter_offence_nas() |>
  filter(offence != "n/a") |>
  filter(mode == "cyclist") |>
  count(offence, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_offence)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  mutate(
    offence = ifelse(n < 20, "other", offence)
  ) |>
  group_by(offence) |>
  summarise(n = sum(n), n_hybrid = n()) |>
  arrange(n_hybrid, desc(n)) |>
  select(-n_hybrid) |>
  mutate(`% of total` = scales::percent(n / sum(n), accuracy = 0.1)) |>
  rename_all(snakecase::to_title_case) |> 
  knitr::kable()
```

In terms 'disposal', the most common values are shown in @tbl-disposal.

```{r}
#| label: tbl-disposal
#| tbl-cap: "Most common disposal values in the OpSnap dataset."
d_all |>
  count(disposal, sort = TRUE) |>
  mutate(percent_records = n / nrow(d_all)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  arrange(desc(n)) |>
  rename_all(snakecase::to_title_case) |> 
  knitr::kable()
```

There are `r unique(d$location) |> length()` unique location text strings (addresses) in the data, with the most common locations shown in @tbl-locations:

```{r}
#| echo: false
#| label: tbl-locations
#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
d |>
  count(location, sort = TRUE) |>
  mutate(percent_records = n / nrow(d)) |>
  mutate(percent_records = scales::percent(round(percent_records, 3))) |>
  arrange(desc(n)) |>
  head(10) |>
  rename_all(snakecase::to_title_case) |> 
  knitr::kable()
```
